\section{Dark Radiation during BBN}
\begin{enumerate}[label=\alph*)]
   \item We know the temperature of deuterium production 
      \begin{equation*}
         T_{NS} \approx \SI{80}{\kilo\eV}
      \end{equation*}
      In order to compute the age of the Universe, need $g_*$ first. Neutrinos decoupled at $\sim \SI{2}{\mega\eV}$ and their temeratures are $(4/11)^{1/3}$. Then
      \begin{equation}
         g_* = 2 + \frac{7}{8} \cdot 3 \cdot 2 \cdot \left( \frac{4}{11} \right)^{4/3} = 3.36
      \end{equation}
      As a result,
      \begin{equation}
         t_{NS} = \frac{M_{Pl}}{1.66\sqrt{g_*}} \frac{1}{2 T^2_{NS}} = \SI{203.3}{\s}
      \end{equation}
   \item Neutron freeze-out must happened before the $T_{NS}$, otherwise no way to burn the deuterium to produce Helium-4. The $g_*$ is then the same as last part.
      \begin{align}
         \Gamma_n &\stackrel{!}{=} H \notag \\
         C_n G_F^2 T^5 &\stackrel{!}{=} T^2 / M_{Pl}^* \notag \\
         T_n^3 &= \left(M_{Pl}^* C_n G_F^2 \right)^{-1} \notag \\
         T_n &= \left(M_{Pl}^* C_n G_F^2 \right)^{-1/3} = \SI{1.16}{\mega\eV}
      \end{align}
   \item
      If there were yet-to-be-discovered light particle, which could be in thermal equilirum with photons in present era, the freeze-out temperature will also be different, since we have explicitely used $g_*$ in the previous calculation. 

      Specifically, $g_*$ would be larger and consequently $T_n \propto g_*^{1/6}$ will be higher and the neutron decoupling would have happened slightly earlier. Since at this temperature neutrons and protons are non-relativistic, they obey Boltzmann distribution $n_{n,p}(T) \propto T_n^{3/2} e^{-m/T}$. For the sake of simplicity, we just ignore the mass difference between protons and neutrons here.       

      In the expression of $\nnnp$, we see $t_{NS}$ and it depends on $g_*$ in the sense that $g_*$ grows, $t_{NS}$ decreases. Then the $\nnnp$ will gets larger and in the end, Helium-4 will become more abundant.

   \item 
      First express effective degrees of freedom in terms of $\Delta N_{\text{eff}}$
      \begin{equation}
         g_* = 2 + \frac{7}{8} \cdot (3 + \Delta N_\text{eff}) \cdot 2 \cdot \left( \frac{4}{11} \right)^{4/3}
      \end{equation}
      
      $g_*$ dependence enters not only $T_n$, but also $t_{NS}$. Thus
      \begin{align*}
         \dv{X_{\fHe}}{N_\eff} &= - \frac{2}{ \left(1 + \frac{n_p(T_{NS})}{n_n(T_{NS})} \right)^2  } \dv{\frac{n_p(T_{NS})}{n_n(T_{NS})}}{N_\eff} \\
                               &= - \frac{X_{\fHe}^2}{2} \left( \pdv{\npnn}{T_n} \pdv{T_n}{N_\eff} + \pdv{\npnn}{t_{NS}}\pdv{t_{NS}}{N_\eff} \right)
      \end{align*}
      Write $\dd$ into $\Delta$ and rearrange the equation, we have
      \begin{equation}
         \frac{\Delta X_\fHe}{X_\fHe} = - \frac{X_\fHe}{2} \left( \pdv{\npnn}{T_n} \pdv{T_n}{N_\eff} + \pdv{\npnn}{t_{NS}}\pdv{t_{NS}}{N_\eff} \right) \Delta N_\eff
      \end{equation}
      It contains four derivatives, thus it is more economical to calculate it with Mathematica. With $T_n \approx \SI{1.4}{\mega\eV}$, $t_{NS}\approx \SI{200}{\s}$, $X_\fHe \approx 45\% $, $\Delta N_eff=1$, and $\mu_n = \mu_p$, we get 
      \begin{equation}
         \frac{\Delta X_\fHe}{X_\fHe} \approx 2.8\%
      \end{equation}
\end{enumerate}
